---
title: "shiny"
author: "snowbcat"
date: "12/9/2021"
output: html_document
---

```{r packages, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(sf) #spatial data
library(tigris) #geojoin
library(leaflet) #interaction map
library(htmlwidgets) #interaction map labels
library(dplyr)
```



## Number of Partially/Fully Vaccinated Individuals in Each Continent for 2021 
```{r read in data}
Vaccination_Across_Months_2_ <- read_excel("Vaccination_Across_Months (2).xlsx")
```

Vaccination_Across_Months_2_ <- read_excel("Vaccination_Across_Months (2).xlsx") %>%
  leaflet() %>%
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(label = Vaccination_Across_Months_2_,
              stroke = FALSE,
              smoothFactor = 0.5,
              opacity = 1, 
              fillOpacity = 0.7, 
              fillColor = ~ pal(caserate),
              highlightOptions = highlightOptions(weight = 5, 
                                                  fillOpacity = 1,
                                                  color = "black",
                                                  opacity = 1, 
                                                  bringToFront = TRUE))
